---
title: "Diamonds_and_Price_Predictions"
author: "Aniko Lengyel"
output: html_document
---


```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

library(ggplot2)
library(GGally)
library(scales)
library(memisc)
library(gridExtra)
library(RColorBrewer)
```

I loaded the data and looked at the structure.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

data("diamonds")

str(diamonds)

```


```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

head(diamonds)

```

"In this case, we can use a variety of diamond characteristics to help us figure
out whether the price advertised for any given diamond is reasonable or a 
rip-off."

I used scatteplots to examine two variables, price(y) and carat_weight(x)
I limited the x-axis and y-axis to exclude the top 1% of values.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

ggplot(aes(x = carat, y = price), data = diamonds) +
  geom_point(alpha = 1/10, fill = I('#9ac5db'), color = I('blue'))+
  scale_x_continuous(limits = c(0, quantile(diamonds$carat, 0.99)))+
  scale_y_continuous(limits = c(0, quantile(diamonds$price, 0.99))) +
  stat_smooth(method = 'lm')

```

I created two plots, following the instructions:

"Create two histograms of the price variable and place them side by side on 
one output image.

We've put some code below to get you started.

The first plot should be a histogram of price and the second plot should 
transform the price variable using log10.""

I created appropriate bin widths for each plot and added a title and filled 
the plots with color.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

plot1 <- qplot(x = price, data = diamonds, bins = 30, fill = I('#9ac5db')) + 
  ggtitle('Price')

plot2 <- qplot(x = price, data = diamonds, bins = 100, fill = I('#9ac5db')) +
  ggtitle('Price (log10)') +
  scale_x_log10()

grid.arrange(plot1, plot2)

```

The function cuberoot_trans transforms the x-axis. Transform function 
takes a cuberoot of a variable, inverse function undoes the results of the 
transform function. We also added limits, breaks and title to make the plots better
readable.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

cuberoot_trans = function() trans_new('cuberoot', transform = function(x) x^(1/3),
                                      inverse = function(x) x^3)

ggplot(aes(carat, price), data = diamonds) + 
  geom_point() + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')

```

I added a color layer to grop the diamonds by clarity.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

ggplot(aes(x = carat, y = price), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter', aes(color = clarity)) +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Clarity', reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')

```

Another color layer to mark the diamonds by cut.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

ggplot(aes(x = carat, y = price, color = cut), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Cut', reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Cut')

```

Another colot layer to gorp the diamonds by color.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

ggplot(aes(x = carat, y = price, color = color), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Color', reverse = FALSE,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Color')


```

We built a linear model to predict the diamond price based on different variables.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = diamonds)
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1, m2, m3, m4, m5)

```

Making a sample prediction based on our linear model.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

thisDiamond = data.frame(carat = 2.5, cut = "Very Good",
                         color = "H", clarity="VS1")
modelEstimate = predict(m5, newdata = thisDiamond,
                        interval="prediction", level = .95)

```