---
title: "Explore_Many_Variables_R_Udacity"
author: "Aniko Lengyel"
output: html_document
---

During my tasks, I analyzed Udacity's pseudo-Facebook data based on the course
instructions. I used an rmd file to show and track the whole analyzis process.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

library(ggplot2)
library(gridExtra)
library(GGally)
library(doBy)
library(corrplot)
```

I red in the dataset and looked at the strucure.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
names(pf)

```

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

head(pf)

```

I created a new data frame, called 'pf.fc_by_age_gender', that 
contains information on each age and gender group, including the following 
variables:

- mean_friend_count,
- median_friend_count,
- n (the number of users in each age and gender grouping)


```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

users_by_age_and_gender <- group_by(pf, age, gender)
pf.fc_by_age_gender<-summarise(users_by_age_and_gender,
                               mean_friend_count = mean(friend_count),
                               median_friend_count = median(friend_count),
                               n = n())

names(users_by_age_and_gender)


```

 I created a line graph to show the median friend count over the ages for each 
 gender in the pf.fc_by_age_gender data frame.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

ggplot(aes(x = age, y = median_friend_count), 
       data = subset(pf.fc_by_age_gender, !is.na(gender)))+
  geom_line(aes(color = gender), stat = "summary")

```

Using package "tidyr".


```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

spread(subset(pf.fc_by_age_gender, 
              select = c('gender', 'age', 'median_friend_count')), 
       gender, median_friend_count)

```

I reshaped the data with dcast function, and excluded the records where there is 
no data about the genders. I also created a new dataframe, named 
pf.fc_by_age_gender.wide.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

pf.fc_by_age_gender.wide <- dcast(subset(pf.fc_by_age_gender, !is.na(gender)), 
                                  age ~ gender,
                                  value.var = 'median_friend_count')

head(pf.fc_by_age_gender.wide)


```

I plotted the ratio of the female to male median friend counts and added a 
horizontal line to the plot with a y intercept and made it dashed. I arranged
the plots to get two diagram based on gender.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

ggplot(aes(x = age, y = female/ male), 
       data = pf.fc_by_age_gender.wide)+
  geom_hline(yintercept=1, linetype = 2, alpha = 0.3)+
  geom_line()

female_plot <- ggplot(aes(x = age, y = female), data = pf.fc_by_age_gender.wide)+
  geom_hline(yintercept=1, linetype = 2)+
  geom_line(stat = "summary")

male_plot <- ggplot(aes(x = age, y = male), data = pf.fc_by_age_gender.wide)+
  geom_hline(yintercept=1, linetype = 2)+
  geom_line

grid.arrange(female_plot, male_plot)

```

I created a new variable, called year_joined in the pf data frame using the variable
tenure. 2014 is the last year.

According to the instructions, the variable year_joined had to contain the year 
that a user joined facebook.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

pf$year_joined <- floor(2014 - pf$tenure/365)

summary(pf$year_joined)

```

Doublecheking the numbers of users joined in each year.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

table(pf$year_joined)

```

I created a new variable in the data frame called year_joined.bucket by using
the cut function on the variable year_joined.

I had to create the following buckets for the new variable, year_joined.bucket.

        (2004, 2009]
        (2009, 2011]
        (2011, 2012]
        (2012, 2014]

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

pf$year_joined.bucket <- cut(pf$year_joined, c(2004, 2009, 2011, 2012, 2014))


```

I had to create a line graph of friend_count vs. age so that each year_joined.bucket 
is a line tracking the median user friend_count across age.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

ggplot(aes(x = age, y = friend_count), 
       data = subset(pf, !is.na(year_joined)))+
  geom_line(aes(color = year_joined.bucket), stat = "summary", fun.y = median)

```

After that I followed the instructions below:

(1) Add another geom_line to code below to plot the grand mean of the friend 
count vs age.

(2) Exclude any users whose year_joined.bucket is NA.

(3) Use a different line type for the grand mean.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

ggplot(aes(x = age, y = friend_count), 
       data = subset(pf, !is.na(year_joined)))+
  geom_line(aes(color = year_joined.bucket), stat = "summary", fun.y = mean)+
  geom_line(stat = 'summary', fun.y = mean, linetype = 2)

with(subset(pf, tenure >= 1), summary(friend_count / tenure))

```

I created a graph of mean of friendships_initiated per day (of tenure)
vs. tenure colored by year_joined.bucket. I exluded the users with less than
one day of tenure.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

ggplot(aes(x = tenure, y = friendships_initiated / tenure), 
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket))

```

I limited the y axis to get a better readable diagram.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

ggplot(aes(x = tenure, y = friendships_initiated / tenure), 
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket)) +
              scale_y_continuous(limits = c(0, 25))

```

I used geom_smooth() to add a smoother line to the plot.
You can use the defaults for geom_smooth() but do color the line
by year_joined.bucket. I named the x, y axis and gave a title.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}

final_plot <- ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_smooth(aes(color = year_joined.bucket)) +
  xlab('Tenure (Day)') +
  ylab('Frienships Initiated') +
  ggtitle('Friendships Initiated per Day (Tenure)')

```

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}



```

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}



```